---
title: "Testing Statistical Graphics"
author: "Susan Vanderplas"
format: 
  revealjs:
    navigation-mode: vertical
    chalkboard: true
    includes:
      in_header: |
        <style>
        @import url('https://fonts.googleapis.com/css2?family=Handlee&family=Montserrat:ital,wght@0,400;0,500;1,400;1,500&family=Roboto:ital,wght@0,500;1,500&display=swap');
        </style>
    theme: [styles.css, fonts.css, default]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev.args = list(png = list(type = "cairo", bg="transparent")))
```

# Introduction

## Why Test Graphics? {background-color="#001226"}

```{r HR-diagram}
#| message: false
#| warning: false
#| echo: false
#| fig-width: 8
#| fig-height: 5
#| out-width: "100%"
#| fig-format: png

library(tidyverse)

# download.file("https://github.com/astronexus/HYG-Database/blob/master/hygfull.csv?raw=true", destfile = "stars.csv")
stars <- readr::read_csv("../data/stars.csv")

stars <- stars %>%
  mutate(Spectral.Class = str_extract(Spectrum, "^.") %>%
           str_to_upper() %>%
           factor(levels = c("O", "B", "A", "F", "G", "K", "M"), ordered = T),
         EarlyLate = str_extract(Spectrum, ".(\\d)") %>%
           str_replace_all("[A-z]", "") %>% as.numeric(),
         Temp = 4600*(1/(.92*ColorIndex + 1.7) + 1/(.92*ColorIndex) + 0.62)) %>%
  filter(!is.na(Spectral.Class) & !is.na(EarlyLate) & !is.na(Hip)) %>%
  arrange(Spectral.Class, EarlyLate) %>%
  mutate(SpectralClass2 = paste0(Spectral.Class, EarlyLate) %>% factor(., levels = unique(.)))

ggplot(data = filter(stars, Distance < 500)) + 
  # annotate(x = -.25, xend = .75, y = -2, yend = -6.5, arrow = arrow(ends = "both", length = unit(.1, "cm")), geom = "segment", color = "grey") + 
  ggplot2::annotate(x = 0.125, xend = 2, y = 4.25, yend = 4.25, arrow = arrow(ends = "both", length = unit(.1, "cm")), geom = "segment", color = "grey") + 
  geom_point(aes(x = ColorIndex, y = -AbsMag, color = Spectral.Class), alpha = .5) + 
  scale_x_continuous("B-V Color Index", breaks = c(0, .5, 1, 1.5, 2), labels = c("Hot  0.0       ", "0.5", "1.0", "1.5", "           2.0  Cool")) + 
  scale_y_continuous("Absolute Magnitude (Brightness)", breaks = c(-8, -4, 0, 4), labels = c(8, 4, 0, -4)) + 
  scale_color_manual("Spectral\nClass", values = c("#2E478C", "#426DB9", "#B5D7E3", "white", "#FAF685", "#E79027", "#DA281F")) + 
  ggplot2::annotate(x = .25, y = -5.5, label = "Dwarfs", geom = "text", angle = -25, color = "white") + 
  ggplot2::annotate(x = .5, y = -3.75, label = "Main Sequence", geom = "text", angle = -28) + 
  ggplot2::annotate(x = 1.125, y = 0, label = "Giants", geom = "text") + 
  ggplot2::annotate(x = 1, y = 4.5, label = "Supergiants", geom = "text", color = "white") +
  theme(panel.background = element_blank(), legend.key = element_blank(), 
        text = element_text(size = 18, color = "white"), 
        plot.background = element_blank()) + 
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "transparent",colour = NA),
    plot.background = element_rect(fill = "transparent",colour = NA),
    legend.background = element_rect(fill = "transparent", color = NA)
  ) +
  ggtitle("Hertzsprung-Russell Diagram") + 
  coord_cartesian(xlim = c(-.25, 2.25), ylim = c(-12, 7)) + 
  guides(color = guide_legend(override.aes = list(alpha = 1)))
```

## Testing Graphics to Communicate Well

::: r-fit-text
Some graphics "just work" - why do you need to test them?

-   Do they work for everyone?

-   What insights do people gain from the chart? How much variability is there?

-   What background knowledge is required?

-   What areas of the chart contribute to its' function?
:::

## Testing Graphics to Understand Data

::: r-fit-text
Testing graphics provides insight into the interaction between humans and data

-   Graphical hypothesis testing with lineups

-   Visual Statistics

-   Effective Science Communication:

    -   Risk (public health, hurricanes, climate)
    -   Legal Evidence weight
    -   Information dissemination
:::

## Testing Graphics to Understand Statistics

::: r-fit-text
Testing graphics can lead to improved methods

-   visualizations for method diagnostics\
    feed back into numerical diagnostics

-   designing algorithms that mimic human perception

-   examining how people understand statistics\
    explain new methods in better ways
:::

## Why Not Test Graphics?

::: r-fit-text
-   People are a PITA (ðŸ’¥âž¡ï¸ðŸ‘)

-   Data collection takes time/energy/money

-   Analyzing (usually very messy) human data is hard

[All of these are true, but it's still worth doing.]{.emph .cerulean .center}
:::

# Discussion

## Biggest Challenges In Testing Graphics

## Most Interesting (Unexpected) Findings

## Things You Wish You Knew When You Started

## Favorite Tools for Graphics Experiments

## Things You Want To Test Next?
